---
title: "ERA 5 Gap Filling quality Assessment"
output: html_notebook
---

```{r, include = F}
library(tidyverse)
theme_set(theme_bw())
library(Metrics)
library(lubridate)
library(xfun)
devtools::load_all()
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
meteo_vars <-
  list(
    "Tair" = "TA_F",
    "SW_IN" = "SW_IN_F",
    "LW_IN" = "LW_IN_F" ,
    "VPD" = "VPD_F",
    "P" = "P_F",
    "PA" = "PA_F"
  )

meteo_vars_eta <- names(meteo_vars) %>%
        paste0("_ERA")

hai_era_path <- here::here("data/FLX_DE-Hai_FLUXNET2015_ERAI_HH_1989-2014_1-4.csv")
hai_path <- here::here("data/FLX_DE-Hai_FLUXNET2015_FULLSET_HH_2000-2012_1-4.csv")
site_raw <- read_csv(hai_path, col_types = cols(TIMESTAMP_START = col_datetime(format="%Y%m%d%H%M"), TIMESTAMP_END = col_datetime(format="%Y%m%d%H%M")), na = c("-9999", "-9999.99"))

# Remove gap filled measurements
site_obs <- site_raw %>%
        filter((if_all(as.character(meteo_vars), \(x) {
          name <- cur_column()
          cur_data()[str_glue("{name}_QC")] == 0
        }))) %>%
        select(TIMESTAMP_END, !!!meteo_vars)

site_era <- read_csv(hai_era_path, col_types = cols(TIMESTAMP_START = col_datetime(format="%Y%m%d%H%M"), TIMESTAMP_END = col_datetime(format="%Y%m%d%H%M")), na = c("-9999", "-9999.99")) %>%
        rename(Tair_ERA = TA_ERA) %>%
        select(TIMESTAMP_END, !!!meteo_vars_eta)

site <- site_obs %>%
        left_join(site_era, by = "TIMESTAMP_END")
```

```{r}
# calculate RMSE for each variable
walk(seq_along(meteo_vars), function(i){
  var_name <- sym(names(meteo_vars)[[i]])
  var_name_era <- sym(paste0(var_name, "_ERA"))
  site <<- site %>%
          mutate(
                  "{var_name}_err" := !!var_name - !!var_name_era
          )
})

# more tidyversy version, but a bit slower
# site %>%
#   mutate(across(
#     names(meteo_vars),
#     \(x) {
#       var_name <- cur_column()
#       x - cur_data()[paste0(var_name, "_ERA")]
#     },
#     .names = "{.col}_err"
#   ))

site_stat <- site %>%
        summarize(
          across(ends_with("err"),
              list( rmse = ~sqrt(mean(.x ^ 2)))
        ))
```

# Introduction

The data used is the downscaling of the ERA-Interim dataset done by FLUXNET.

It includes a correction for local conditions using a linear model and linear interpolation for the reducing downscaling the time measurement.

# Analysis



## Tair

The air temperature RMSE is about 1.9 ° C and overall the model is quite good.

The scatter plot is good and by analysing the timeseries the ERA dataset models accurately the overall trend in temperature but misses on the finer scale variation.

In conclusion, while there is still some room for improvement the gap filling is quite good

```{r, results = 'asis', message = FALSE, echo=FALSE}
knit_template("analysis/fragments/variable_assess_ERA_gapfilling.rmd", var = "Tair", unit="°C")
```

## SW_IN

The RMSE of the SW_is better, but still has a lot of variation.

This is likely due to the different cloud coverage between local conditions and the overall conditions that the ERA-5 dataset cannot model and which lead sometimes to significant errors (erg. gap sample 5)



```{r, results = 'asis', echo = FALSE}
knit_template("analysis/fragments/variable_assess_ERA_gapfilling.rmd", var = "SW_IN", unit="W m-2")
```

## LW_IN

The gap filling for LW_IN is okay but not very good.

One key driver of the LW_IN is the amounts of clouds, which is a local condition and likely the main source of errors for LW_IN



```{r, results = 'asis', echo = FALSE}
knit_template("analysis/fragments/variable_assess_ERA_gapfilling.rmd", var = "LW_IN", unit="W m-2")
```
## VPD


The vapour pressure deficit gap filling is overall good.

The gap filled distribution has a shifted mode and in the sample 4 doesn't follow well the time series

```{r, results = 'asis',  echo = FALSE}
knit_template("analysis/fragments/variable_assess_ERA_gapfilling.rmd", var = "VPD", unit="hPa")
```

## P

The precipitation intensity is usually underestimated in the ERA dataset.

However, precipitation is a very hard variable and probably the comparison should be on sum over a period on for each observation


```{r, results = 'asis',  echo = FALSE}
knit_template("analysis/fragments/variable_assess_ERA_gapfilling.rmd", var = "P", unit="mm")
```


## PA

The air pressure is simply amazing! The average error is less than the instrument accuracy.

This is expected as the local conditions should have a limited impact on the pressure

```{r, results = 'asis',  echo = FALSE}
knit_template("analysis/fragments/variable_assess_ERA_gapfilling.rmd", var = "PA", unit="hPa")
```

# Conclusion

The ERA based gap filling has overall a good performance. In particular it gets right the overall trends of the data, which limits the magnitude of the errors, however often doesn't follow precisely local conditions.

This can be a problem especially when filling small gaps.