---
title: "MDS Multiple Gap Filling Quality assessment"
output: html_notebook
---
```{r, include = FALSE}
library(tidyverse)
theme_set(theme_bw())
library(Metrics)
library(lubridate)
library(xfun)
# parallel execution
library(furrr)
n_workers <- 11
plan(multisession, workers = n_workers)

# number repetition for each gap
n_rep <-  30

#load_all() doesn't works with futures, so loading only code in R dir
list.files(here::here("R"), full.names = TRUE) %>%
  walk(source)
collapsible_hook() # register the collpasible hook

knitr::opts_chunk$set(echo = FALSE, cache.path = str_glue("cache/hai_{n_rep}/"))
```

# Introduction

## Marginal Distribution Sampling (MDS)

Assess the quality of the gap filling using the default algorith from `REddyProc`, a standard implementations of meteo gap filling in the Eddy Covariance.

The variables used for filling gaps are, that are used according to the REddyProc default:

- Shortwave radiation incoming `SW_IN`
- Air Temperature `Tair`
- Vapour pressure deficit `VPD`

## Analysed variables

The variables that are gap filled are:

- Temperature
- Shortwave radiation incoming
- Longwave radiation incoming
- Vapour Pressure Deficit
- Pressure
- Precipitation


## Instruments accuracy

In order to estimate the quality of the gap filling the RMSE is compared with the instrument accuracy for the sensors.

The list of instruments available at Hainich is available at [https://data.icos-cp.eu/objects/_tFsWRgQcO7FkfvOq0OqIC8H](https://data.icos-cp.eu/objects/_tFsWRgQcO7FkfvOq0OqIC8H)

| Variable            | Instrument      | Accuracy | Reference                                                                                  |
|---------------------|-----------------|----------|--------------------------------------------------------------------------------------------|
| Temperature         | Vaisala HMP35   | 0.1 °C   | https://docs.vaisala.com/access/sources/ud/document?aton_code=B211826EN                    |
| Humidity            | Vaisala HMP35   | 0.8% RH  | https://docs.vaisala.com/access/sources/ud/document?aton_code=B211826EN                    |
| Pressure            | Vaisala PTB101B | 0.3 hPa  | https://www.vaisala.com/sites/default/files/documents/PTB110-Datasheet-B210681EN.pdf       |
| Precipitation       | OTT Pluvio2     | 0.1mm    | https://www.ott.com/download/leaflet-compact-weighing-precipitation-gauge-ott-pluvio2-s-1/ |
| Shortwave radiation | DeltaT SPN1     | 10 Wm-2  | https://delta-t.co.uk/wp-content/uploads/2016/10/SPN1-Technical-Fact-Sheet-v1.2_d_web.pdf  |


## Load Data

The data from the Eddy Covariance station in Hainich is used for the analysis.

The data is from the FLUXNET 2015 release

```{r}
hai_path <- here::here("data/FLX_DE-Hai_FLUXNET2015_FULLSET_HH_2000-2012_1-4.csv")
hai_raw <- read_csv(hai_path, col_types = cols(TIMESTAMP_START = col_datetime(format="%Y%m%d%H%M"), TIMESTAMP_END = col_datetime(format="%Y%m%d%H%M")), na = c("-9999", "-9999.99"))
```
```{r}
meteo_vars <-
  list(
    "Tair" = "TA_F",
    "SW_IN" = "SW_IN_F",
    "LW_IN" = "LW_IN_F" ,
    "VPD" = "VPD_F",
    "P" = "P_F",
    "PA" = "PA_F"
  )

meteo_qc <- c(
  "TA_F_QC", "SW_IN_F_QC", "LW_IN_F_QC",  "VPD_F_QC", "P_F_QC", "PA_F_QC"
)
# Note this include the oberservations that are gap filled.
# REddyProc doesn't work if there are gaps in the timeseries so can't remove them from the dataset
hai <- hai_raw %>%
  mutate(
    Tair = na_if(TA_F, TA_F_QC !=0),
    SW_IN = na_if(SW_IN_F, SW_IN_F_QC !=0),
    LW_IN = na_if(LW_IN_F, LW_IN_F_QC != 0),
    VPD = na_if(VPD_F, VPD_F_QC !=0),
    P = na_if(P_F, P_F_QC != 0),
    PA = na_if(PA_F, PA_F_QC != 0)
  ) %>%
  select(TIMESTAMP_END, !!!meteo_vars)

site_data <- hai
```

## RMSE

To have a simple test of the accuracy of the gap filling algorith artificial gaps are created and the Root Mean Square Error (RMSE) is computed.

For each of the variable of interest:

 - **`r n_rep`** gaps for each of 30 different lengths (ranging from 1 hour to 60 days) at random location are generated
 - the gaps are filled using `REddyProc`
 - for the gap period RMSE is calculated
 - for each gap length the RMSE and the standard deviation is calculated

generate artificial gaps and then fill the data using REddyProc.
```{r}
# go up to 60 days which is the maximium gap is going to be filled
# 1.1 is needed because for some unknown reasons, when converting to interger 1 becomes 0 (maybe some floating point related issue)
gaps_lengths <- seq_log(1.1, 24 * 2 * 30 * 2, offset = 15, length.out = 30) %>%
  rep(n_rep) %>%
  as.integer()
# get 4 gap lengths from the gap (escluding the last one)
gap_sample_idx <- seq(from = 2, to = length(unique(gaps_lengths)), length.out = 5) %>% as.integer()
gap_sample <- gaps_lengths[gap_sample_idx[1:4]]
```
```{r}
# run the simulation only for the 4 samples of interest
gaps_lengths_multi <- gap_sample %>%
  rep(n_rep) %>%
  as.integer()
```




# Multiple missing variables




## Tair




```{r, results = 'asis', message = FALSE, echo=FALSE}
knit_template("analysis/fragments/MDS_multiple_missing.rmd", var = "Tair", unit="°C")
```

## SW_IN



```{r, results = 'asis', echo = FALSE}
knit_template("analysis/fragments/MDS_multiple_missing.rmd", var = "SW_IN", unit="W m-2")
```

## LW_IN

```{r, results = 'asis', echo = FALSE}
knit_template("analysis/fragments/MDS_multiple_missing.rmd", var = "LW_IN", unit="W m-2")
```


## VPD



```{r, results = 'asis',  echo = FALSE}
knit_template("analysis/fragments/MDS_multiple_missing.rmd", var = "VPD", unit="hPa")
```

## P



```{r, results = 'asis',  echo = FALSE}
knit_template("analysis/fragments/MDS_multiple_missing.rmd", var = "P", unit="mm")
```


## PA


```{r, results = 'asis',  echo = FALSE}
knit_template("analysis/fragments/MDS_multiple_missing.rmd", var = "PA", unit="hPa")
```


# Conclusion


