---
title: "MDS Gap Filling Quality assessment"
output: html_notebook
---
```{r, include = FALSE}
library(tidyverse)
theme_set(theme_bw())
library(Metrics)
library(lubridate)
library(xfun)
# parallel execution
library(furrr)
n_workers <- 11
plan(multisession, workers = n_workers)

# number repetition for each gap
n_rep <-  50

#load_all() doesn't works with futures, so loading only code in R dir
list.files(here::here("R"), full.names = TRUE) %>%
  walk(source)
```

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, cache.path = str_glue("cache/hai_{n_rep}/"))
```

# Marginal Distribution Sampling (MDS)

Assess the quality of the gap filling using the default algorith from `REddyProc`, a standard implementations of meteo gap filling in the Eddy Covariance.

The variables used for filling gaps are, that are used according to the REddyProc default:

- Shortwave radiation incoming `SW_IN`
- Air Temperature `Tair`
- Vapour pressure deficit `VPD`

## Analysed variables

The variables that are gap filled are:

- Temperature
- Shortwave radiation incoming
- Longwave radiation incoming
- Vapour Pressure Deficit
- Pressure
- Precipitation


## Instruments accuracy

In order to estimate the quality of the gap filling the RMSE is compared with the instrument accuracy for the sensors.

The list of instruments available at Hainich is available at [https://data.icos-cp.eu/objects/_tFsWRgQcO7FkfvOq0OqIC8H](https://data.icos-cp.eu/objects/_tFsWRgQcO7FkfvOq0OqIC8H)

| Variable            | Instrument      | Accuracy | Reference                                                                                  |
|---------------------|-----------------|----------|--------------------------------------------------------------------------------------------|
| Temperature         | Vaisala HMP35   | 0.1 째C   | https://docs.vaisala.com/access/sources/ud/document?aton_code=B211826EN                    |
| Humidity            | Vaisala HMP35   | 0.8% RH  | https://docs.vaisala.com/access/sources/ud/document?aton_code=B211826EN                    |
| Pressure            | Vaisala PTB101B | 0.3 hPa  | https://www.vaisala.com/sites/default/files/documents/PTB110-Datasheet-B210681EN.pdf       |
| Precipitation       | OTT Pluvio2     | 0.1mm    | https://www.ott.com/download/leaflet-compact-weighing-precipitation-gauge-ott-pluvio2-s-1/ |
| Shortwave radiation | DeltaT SPN1     | 10 Wm-2  | https://delta-t.co.uk/wp-content/uploads/2016/10/SPN1-Technical-Fact-Sheet-v1.2_d_web.pdf  |


## Load Data

The data from the Eddy Covariance station in Hainich is used for the analysis.

The data is from the FLUXNET 2015 release

```{r}
hai_path <- here::here("data/FLX_DE-Hai_FLUXNET2015_FULLSET_HH_2000-2012_1-4.csv")
hai_raw <- read_csv(hai_path, col_types = cols(TIMESTAMP_START = col_datetime(format="%Y%m%d%H%M"), TIMESTAMP_END = col_datetime(format="%Y%m%d%H%M")), na = c("-9999", "-9999.99"))
```
```{r}
meteo_vars <-
  list(
    "Tair" = "TA_F",
    "SW_IN" = "SW_IN_F",
    "LW_IN" = "LW_IN_F" ,
    "VPD" = "VPD_F",
    "P" = "P_F",
    "PA" = "PA_F"
  )

meteo_qc <- c(
  "TA_F_QC", "SW_IN_F_QC", "LW_IN_F_QC",  "VPD_F_QC", "P_F_QC", "PA_F_QC"
)
# Note this include the oberservations that are gap filled.
# REddyProc doesn't work if there are gaps in the timeseries so can't remove them from the dataset
hai <- hai_raw %>%
    mutate(
      Tair = na_if(TA_F, TA_F_QC !=0),
      SW_IN = na_if(SW_IN_F, SW_IN_F_QC !=0),
      LW_IN = na_if(LW_IN_F, LW_IN_F_QC != 0),
      VPD = na_if(VPD_F, VPD_F_QC !=0),
      P = na_if(P_F, P_F_QC != 0),
      PA = na_if(PA_F, PA_F_QC != 0)
    ) %>%
    select(TIMESTAMP_END, !!!meteo_vars)

site_data <- hai
```

# RMSE

To have a simple test of the accuracy of the gap filling algorith artificial gaps are created and the Root Mean Square Error (RMSE) is computed.

For each of the variable of interest:

 - **`r n_rep`** gaps for each of 30 different lengths (ranging from 1 hour to 60 days) at random location are generated
 - the gaps are filled using `REddyProc`
 - for the gap period RMSE is calculated
 - for each gap length the RMSE and the standard deviation is calculated

generate artificial gaps and then fill the data using REddyProc

## Tair

The mean RMSE for the temperature ranges between 2 and 5 째C.
This can be considered an high, given that the instrument accuracy is 0.1 째C.
Moreover, temperature is very important variable for the ecosystem, but also has good temporal autocorrelation which should simplify gap filling.


```{r, results = 'asis', message = FALSE, echo=FALSE}
knit_template("analysis/fragments/variable_assess_MDS_gapfilling.Rmd", var = "Tair", unit="째C")
```

## SW_IN

The incoming shortwave radiation average RMSE is 100 W m-2
This is roughly 10 times the instrument sensitivity.
Incoming shortwave behaviour should also be relatively easy to predict.

This should be checked how it varies across the year, the SW_IN values can be quite different between winter and summer

```{r, results = 'asis', echo = FALSE}
knit_template("analysis/fragments/variable_assess_MDS_gapfilling.Rmd", var = "SW_IN", unit="W m-2")
```
## VPD

The instrument accuracy is relative humidity, while the variable of interest is the Vapour Pressure Deficit (VPD)
The VPD is calculated as the difference from the saturation vapour pressure (`e_s`) and the actual vapour pressure(`e_a`).
So to estimate the accuracy of the VPD measurements we are calculating for each observation the accuracy connected of VPD by transforming the RH measurement. Then the VPD accuracy is averaged over
So we are using the formula to estimate the accuracy of the instruments relative to the `VPD`

```{r, echo = T}
get_es <-  function(ta)  6.1078 * exp((17.08085 * ta) / (234.175 + ta))
rh2ea <- function(rh, e_s) rh/100 * e_s
get_vpd <- function(e_s, e_a) e_s - e_a

rh_acc <- .8

vpd_acc <- hai %>%
    mutate(e_s = get_es(Tair),
           e_a_acc = rh2ea(rh_acc, e_s), # .8 is the accuracy of the instrument
           vpd_acc = get_vpd(e_s, e_a_acc)
    )

vpd_acc_mean <- mean(vpd_acc$vpd_acc) %>%  round(2)
```

|              |                      |
|--------------|----------------------|
| VPD Accuracy | `r vpd_acc_mean` hPa |


```{r, include = F, eval = F}
# For reference the distribution of the vpd accuracy
vpd_acc %>%
    ggplot() +
    geom_histogram(aes(vpd_acc), color = "white") +
    geom_vline(aes(xintercept = vpd_acc_mean, colour= "mean"))
```


The average RMSE seems it has a good value, but this is a likely inaccurate due to the high frequency of 0 in the observations

```{r, results = 'asis',  echo = FALSE}
knit_template("analysis/fragments/variable_assess_MDS_gapfilling.Rmd", var = "VPD", unit="hPa")
```

## P

The Precipitation is a hard variable to gap fill, because is erratic and has a limited connection with the other variables.

The precipitation has a high number of 0 this likely is the reason for the good RMSE.

Comparing sums over a period of time may be a better way to assess precipitations than the point-wise RMSE

```{r, results = 'asis',  echo = FALSE}
knit_template("analysis/fragments/variable_assess_MDS_gapfilling.Rmd", var = "P", unit="mm")
```


## PA

The mean RMSE for the Air Pressure ranges between 0.5 and 1 hPa, which is roughly 3 times the instrument accuracy.

```{r, results = 'asis',  echo = FALSE}
knit_template("analysis/fragments/variable_assess_MDS_gapfilling.Rmd", var = "PA", unit="hPa")
```

# Conclusion

The exploratory analysis on the performance of the MDS algorithm for gap filling shows that it is often limited.

