---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
theme_set(theme_bw())
library(Metrics)
library(lubridate)
library(xfun)
# parallel execution
library(furrr)
n_workers <- 12
plan(multisession, workers = n_workers)

# number repetition for each gap
n_rep <-  30

#load_all() doesn't works with futures, so loading only code in R dir
list.files("R", full.names = TRUE) %>%
  walk(source)
```

# Marginal Distribution Sampling (MDS)

Assess the quality of the gap filling using the default algorith from `REddyProc`, one of the standard implementations.

The variables used for filling gaps are, that are used according to the REddyProc default:

- `SW_IN`
- `Tair`
- `VPD`



## Load Data

The data from the Eddy Covariance station in Hainich is used for the analysis.

The data is from the FLUXNET 2015 release
```{r}
hai_path <- here::here("data/FLX_DE-Hai_FLUXNET2015_FULLSET_HH_2000-2012_1-4.csv")
hai_raw <- read_csv(hai_path, col_types = cols(TIMESTAMP_START = col_datetime(format="%Y%m%d%H%M"), TIMESTAMP_END = col_datetime(format="%Y%m%d%H%M")), na = c("-9999", "-9999.99"))
```
```{r}
meteo_vars <-
  list(
    "Tair" = "TA_F",
    "SW_IN" = "SW_IN_F",
    "LW_IN" = "LW_IN_F" ,
    "VPD" = "VPD_F",
    "P" = "P_F",
    "PA" = "PA_F"
  )

meteo_qc <- c(
  "TA_F_QC", "SW_IN_F_QC", "LW_IN_F_QC",  "VPD_F_QC", "P_F_QC", "PA_F_QC"
)
# Note this include the oberservations that are gap filled.
# REddyProc doesn't work if there are gaps in the timeseries so can't remove them from the dataset
hai <- hai_raw %>%
    mutate(
      Tair = na_if(TA_F, TA_F_QC !=0),
      SW_IN = na_if(SW_IN_F, SW_IN_F_QC !=0),
      LW_IN = na_if(LW_IN_F, LW_IN_F_QC != 0),
      VPD = na_if(VPD_F, VPD_F_QC !=0),
      P = na_if(P_F, P_F_QC != 0),
      PA = na_if(PA_F, PA_F_QC != 0)
    ) %>%
    select(TIMESTAMP_END, !!!meteo_vars)
```

generate artificial gaps and then fill the data using REddyProc

## Tair



```{r}
# go up to 60 days which is the maximium gap is going to be filled
# 1.1 is needed because for some unknown reasons, when converting to interger 1 becomes 0 (maybe some floating point related issue)
gaps_lengths <-  seq_log(1.1, 24 * 2 * 30 * 2 , offset = 20, length.out = 30) %>%
        rep(n_rep) %>%
        as.integer()
```
```{r, message = FALSE}
Tair_rmse <- gaps_lengths %>%
        split_vector(n_workers) %>%
        future_map_dfr(~gap_fill_EProc_rmse(hai, "Tair", .x), seed=TRUE)
```
```{r}
Tair_rmse
```

```{r}
Tair_rmse_stat <- Tair_rmse %>%
  group_by(gap_length) %>%
  summarise(mean = mean(rmse), sd = sd(rmse)) %>%
    mutate(error_high = mean + sd, error_low = mean - sd)
```
```{r}
Tair_rmse_stat %>%
  mutate(gap_length = gap_length / 2) %>%
  ggplot(aes(gap_length)) +
  geom_ribbon(aes(ymin = error_low, ymax = error_high), alpha = .7) +
  geom_line(aes(y=mean)) +
  labs(x = "Gap Lengths [hours]", y = "RMSE [°C]")
```
## SW_IN
```{r, results = 'asis'}
knit_template("analysis/fragments/variable_assess_gapfilling.rmd", var = "SW_IN", unit="W m-2")
```

## P
```{r, results = 'asis'}
knit_template("analysis/fragments/variable_assess_gapfilling.rmd", var = "P", unit="mm")
```

## PA
```{r, results = 'asis'}
knit_template("analysis/fragments/variable_assess_gapfilling.rmd", var = "PA", unit="Pa")
```
## VPD
```{r, results = 'asis'}
knit_template("analysis/fragments/variable_assess_gapfilling.rmd", var = "VPD", unit="Pa")
```

## Instruments accuracy

checking the instruments installed in Hainich [https://data.icos-cp.eu/objects/_tFsWRgQcO7FkfvOq0OqIC8H](https://data.icos-cp.eu/objects/_tFsWRgQcO7FkfvOq0OqIC8H)

| Variable            | Instrument      | accuracy | Reference                                                                                  |
|---------------------|-----------------|----------|--------------------------------------------------------------------------------------------|
| Temperature         | Vaisala HMP35   | 0.1 °C   | https://docs.vaisala.com/access/sources/ud/document?aton_code=B211826EN                    |
| Humidity            | Vaisala HMP35   | 0.8% RH  |                                                                                            |
| Pressure            | Vaisala PTB101B | 0.3 hPa  | https://www.vaisala.com/sites/default/files/documents/PTB110-Datasheet-B210681EN.pdf       |
| Precipitation       | OTT Pluvio2     | 1mm      | https://www.ott.com/download/leaflet-compact-weighing-precipitation-gauge-ott-pluvio2-s-1/ |
| Shortwave radiation | DeltaT SPN1     | 10 Wm-2  | https://delta-t.co.uk/wp-content/uploads/2016/10/SPN1-Technical-Fact-Sheet-v1.2_d_web.pdf  |
