---
title: "REddYProc Testing"
output: html_notebook
---

REddyProc is a standard implementation of the gap filling algorithm for fluxes and meteo data.

in this notebook we are trying to understand how it works and how can be used

```{r}
library(tidyverse)
theme_set(theme_bw())
library(Metrics)
```

# Example

this is the example from their Readme
```{r}


library(REddyProc)
#+++ Input data from csv (example needs to be downloaded)
examplePath <- getExamplePath('Example_DETha98.txt', isTryDownload = TRUE)
if (length(examplePath)) {
  EddyData <- fLoadTXTIntoDataframe(examplePath)
} else {
  warning(
      "Could not find example text data file."
      ," In order to execute this example code,"
      ," please, allow downloading it from github. "
      ," Type '?getExamplePath' for more information.")
  # using RData version distributed with the package instead
  EddyData <- Example_DETha98
}
#+++ If not provided, calculate VPD from Tair and rH
EddyData$VPD <- fCalcVPDfromRHandTair(EddyData$rH, EddyData$Tair)
#+++ Add time stamp in POSIX time format
EddyDataWithPosix <- EddyData %>%
  filterLongRuns("NEE") %>%
  fConvertTimeToPosix('YDH', Year = 'Year', Day = 'DoY', Hour = 'Hour')
#+++ Initalize R5 reference class sEddyProc for processing of eddy data
#+++ with all variables needed for processing later
EProc <- sEddyProc$new(
  'DE-Tha', EddyDataWithPosix, c('NEE','Rg','Tair','VPD', 'Ustar'))
#Location of DE-Tharandt
EProc$sSetLocationInfo(LatDeg = 51.0, LongDeg = 13.6, TimeZoneHour = 1)
#
#++ Fill NEE gaps with MDS gap filling algorithm (without prior ustar filtering)
EProc$sMDSGapFill('NEE', FillAll = FALSE)
#
#++ Export gap filled and partitioned data to standard data frame
FilledEddyData <- EProc$sExportResults()
#
#++ Example plots of filled data to screen or to directory \plots
EProc$sPlotFingerprintY('NEE_f', Year = 1998)
```
# FLuxnet data
```{r}
hai_path <- here::here("../data/FLX_DE-Hai_FLUXNET2015_FULLSET_HH_2000-2012_1-4.csv")
```
```{r}
hai <- read_csv(hai_path, col_types = cols(TIMESTAMP_START = col_datetime(format="%Y%m%d%H%M"), TIMESTAMP_END = col_datetime(format="%Y%m%d%H%M")), na = c("-9999", "-9999.99"))
```
should consider removing gap filled data from here


```{r}
meteo_vars <-
  list(
    "Tair" = "TA_F",
    "SW_IN" = "SW_IN_F",
    "LW_IN" = "LW_IN_F" ,
    "VPD" = "VPD_F",
    "P" = "P_F",
    "PA" = "PA_F"
  )

meteo
hai <- hai %>%
  select(TIMESTAMP_END, !!!meteo_vars)
```
```{r}
EProc_hai <- sEddyProc$new("de-HAI", Data = hai, ColNames = names(meteo_vars), ColPOSIXTime = "TIMESTAMP_END")
```
```{r}
# This takes long 
#EProc_hai$sMDSGapFill("Tair", V1 = "SW_IN", V2 = "VPD", V3 = "Tair")
```
```{r}
EProc_hai$sExportResults()
```

```{r}
227952 / 100
```
```{r}
create_gap <- function(data, var, gap_length){
  gap <- random_gap(gap_length, nrow(data))
  data[gap, var] <- NA
  data
}
```
```{r}
random_gap <- function(gap_length, total_length) {
  # constructing an array to indetify gaps
  # it should be always false (no gap) except for a continous section of lenght `gap_length`
  if(gap_length >= total_length){
    return(rep(1, total_length))
  }
  gap_length <- ifelse(gap_length > total_length, total_length, gap_length)
  gap_start <- sample.int(total_length - gap_length, 1) # gap cannot start too close to the end
  c(
          rep(0, gap_start),
          rep(1, gap_length),
          rep(0, total_length - (gap_length + gap_start))
  )
}
```


In order to have REddyProc understand the gaps properly an additional column with a mask of where the gaps are is needed.
Then this is the argument of `QFVar` and then value of the mask for which there is **no** gap is then passed as `QFValue`



```{r}
fill_gaps_eddy_proc <- function(data, var, id = "ID"){
  EProc <- sEddyProc$new(id, Data = data, ColNames = names(select(data, -TIMESTAMP_END)), ColPOSIXTime = "TIMESTAMP_END")
  EProc$sMDSGapFill(var, QFVar = "gap", QFValue = 0, V1 = "SW_IN", V2 = "VPD", V3 = "Tair", FillAll = FALSE)
  EProc$sExportResults()
}
```

```{r}
hai %>%
  mutate(gap = random_gap(100, nrow(.))) %>%
  fill_gaps_eddy_proc("Tair")
```

```{r}
gap_fill_EProc_rmse_single <- function(data, var, gap_length){
   data_gap <- data %>%
      mutate(gap = random_gap(gap_length , nrow(.)))
  filled <- data_gap %>%
      fill_gaps_eddy_proc(var)
  # some rlang magic needed to have filter() working properly (as it doesn't work with string variable names) with var being a string
  var_fqc <- sym(str_glue("{var}_fqc"))
  var_f <- sym(str_glue("{var}_f"))

  var_filled <- filled %>%
      filter(!!var_fqc != 0) %>%
      pull(!!var_f)

  var_orig <- data_gap %>%
      filter(gap != 0) %>%
      pull(!!var)

  tibble(
      rmse = rmse(var_orig, var_filled),
      gap_length = gap_length
  )
}

#' vectorized version
gap_fill_EProc_rmse <- function(data, var, gap_lengths){
  map_dfr(gap_lengths, ~gap_fill_EProc_rmse_single(data, var, .x))
}
```



```{r}
gap_fill_EProc_rmse_single(hai, "Tair", 1000)
```

## Test different gaps

```{r}
gaps_lengths <- exp(seq(log(3), log(24*2*60), length.out = 30)) %>% # log scale for different lenghts
        rep(100) %>%
        as.integer()
rmse <- gaps_lengths %>% 
  map_dfr(~gap_RMSE(hai, "Tair",.x))
```

```{r}
rmse
```

```{r}
rmse_stat <- rmse %>%
  group_by(gap_length) %>%
  summarise(mean = mean(rmse), sd = sd(rmse))

```

```{r}
rmse_stat %>%
  filter
```

```{r}
rmse_stat %>%
  ggplot(aes(gap_length, mean)) +
  geom_line()

```




```{r}
rmse %>%
  ggplot(aes(gap_length, rmse)) +
  geom_point()

```

```{r}
gaps_lengths[1:30]
```

```{r}
rmse %>%
  filter(!gap_length %in% c(4,6,7,15,19)) %>% #remove some values for better plotting
  mutate(gaps_length = factor(gap_length)) %>%
  ggplot(aes(gap_length, rmse, group = gap_length)) +
  geom_violin()
```

# Tests

```{r}
library(testthat)
```
```{r}
test_that("total lenght",{
          expect_equal(length(get_random_gap(10, 100)), 100)
}

)
```

